Speed estimation for vehicles using \\SVM classification and edge detection

### 2016-06-15 Administratie/report/paper.tex
update paper

\subsection{Precision, recall, F-scores}

TODO: Bjorn

@@

\section{Method}
In order to determine the maximum speed of a vehicle two image processing techniques are used: SVM classification and edge detection. We first try to rely on edge detection to find the contours of the lanes where driving is permitted (i.e. where no obstacles are found). We estimate the reliability of the edge detection and incorporate road detection using the SVM classifier when needed.

For the implementation of our method, we use OpenCV 3.1 \cite{Bradski2000} and C++. Four different datasets were used to either train or test the solution.
To train out SVM classifier, we used a function of the OpenCV SVM library which selects the optimal parameters using cross-validation.

\subsection{Parameters for SVM classification}\label{parameters-classification}
The goal of this classifier is to determine what pixels are part of the road and which are not. We will describe multiple tunable parameters in order to get optimal classification for road detection. These parameters determine both the length and values of the feature vectors, which are then fed to the SVM classifier. We compare classification results varying these parameters for the feature vector in section \ref{results-svm-classification}.

@@

Detecting edges is very straightforward using an edge detection algorithm like Canny Edge Detection \cite{canny}. However, there are a lot of edges on the road itself, thinking of road markings and cobbles. The edges on the road, considered as noise, and the edges of the road are not unambiguously separable. The width of the patterns that cause the edges on the road is rather small. Using morphological image processing operations, it is possible to remove these small patterns.  In figure \ref{cobbles_erode_out}, a smoothed cobbled road is shown.
\begin{figure}[ht]
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\textwidth]{fig/cobbles_in.png}
  \caption{Original frame\label{cobbles_in}}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\textwidth]{fig/cobbles_erode_out.png}
  \caption{Eroded cobbled road\label{cobbles_erode_out}}
\end{subfigure}
\caption{Perform eroding to remove the edges on the road.}
\end{figure}

Eroding is a morphology operator to make the objects on the foreground, which are the brightest, smaller \cite{sreedhar2012enhancement}. Figure \ref{line_zoom_eroded} shows an example of a thinned white line. When the kernel is large enough, white lines can be filtered out completely. The same technique is used to smooth the surface of cobbled roads. A cobblestone consist of a bright center, surrounded by darker joints. By eroding the cobbled road, the erode function will minimize the center of the cobble, which will cause enlarged joints. With a large kernel, the joints will spread until they overlap. The best results were achieved with an ellipsoidal kernel of 100 x 30 px.

@@

Unfortunately, the erode function will cause the darker objects to expand. The original size of the objects has been modified, as seen in \ref{car_erode_bigger}. This will falsify the eventual edge detection. In this example, the observed car is reported to close. In order to restore the original measures, dilation is executed. This is the opposite morphology operator of eroding. The combination of eroding and dilation is called opening. When dilation is executed with the same kernel size of the erode function, the original size of the dark objects will decrease to their original size, as seen in \ref{car_dilate_smaller}. Note that, after eroding, some details of the shape of the objects are lost. In order to provide a safety margin, the dilate function is executed with a slightly smaller kernel.

@@

This algorithm requires a minimum and a maximum threshold value, which were determined experimentally. 
The best results were achieved with a maximum threshold that is twice the minimum threshold.
The minimum threshold is determined by the required minimum sensitivity. When the algorithm is configured too sensitive, some imperfections on the road will be detected as edges. On the other hand, a certain level of sensitivity is necessary to detect less clear road edges. For example, a transition from road to dirt has a very small distinction of color and brightness, as seen in \ref{transition_road_dirt_eroded}. The minimum threshold should be high enough to detect this transition as edge.

@@

\subsection{Zebra crossings}
TODO

@@

For each frame the given masks show the projected trajectory of the car. Intersections get calculated in order to find the road where it's safe to drive. The edges of the edge detection are visualized as blue lines on figure \ref{visualization}. The intersections are shown in green. The closest intersection to the car is shown in red. At this point the maximum speed is derived from the intensity of this pixel.

In order to calculate the edges, both road classification and edge detection are combined. Therefor two strategies were tested:
\begin{enumerate}
\item  A region is considered to be road only if it's identified as road by the road classification and there are no edges. Both classification and edge detection should indicate a pixel as road in order to be used for safe driving.
\item A region is considered to be road if there are no edges. When edges are detected, road classification is used to decide what part of the image is road. 
\end{enumerate}

Both strategies start by removing outliers from the road classification. The goal of this is to eliminate the effect an outlier would have when it's close to the car. A non-road block is considered an outlier when it has less than 4 non-road block neighbours. This is repeated until all outliers are removed.

For the first strategy the edges of the classification are computed. These get combined with the edges from the edge detection. The closest intersection of any edge with the given mask of the image is used to calculate the maximum speed of the vehicle.

In the second strategy, the edges are checked if they aren't on the road. This is done by checking that an edge is on a non-road block of the road classification. For safety reasons there is an additional check. The above, left and right blocks of an edge should also be considered as non-road. When the edge is not classified as road, it is added to a collection. The intersections of the edges in the collection and the masks are then computed.

@@

todo: geen schatten van edge detection meer

@@

TODO: maximum met edge detection voor elke dataset? 1350,

@@

The second strategy for combination also results in zero crashes over the four datasets. This strategy tries to focus on edge detection. A region is considered road if there are no edges. When there are edges, road classification is used to decide whether it's road or not. This strategy gave us the best results, as shown in figure \ref{distance_result} as the fourth bar.
\begin{figure}[ht]
	\centering
	\includegraphics[width=\textwidth]{fig/distance_result.png}
	\caption{Results of road classification, edge detection and combinations.\label{distance_result}}
\end{figure}

>>>>

\subsection{Precision, recall, F-scores}
TODO: Bjorn

@@

\section{Method}
In order to determine the maximum speed of a vehicle two image processing techniques are used: SVM classification and edge detection. In order to combine the best of both techniques, two strategies are discussed in section \ref{method-combination}.

For the implementation of our method, we use OpenCV 3.1 \cite{Bradski2000} and C++. Four different datasets were used to either train or test the solution.

\subsection{Parameters for SVM classification}\label{parameters-classification}
The goal of this classifier is to determine what pixels are part of the road and which are not. We will describe multiple tunable parameters in order to get optimal classification for road detection. These parameters determine both the length and values of the feature vectors, which are then fed to the SVM classifier. We compare classification results varying these parameters for the feature vector in section \ref{results-svm-classification}. To train our SVM classifier, we use a function of the OpenCV SVM library which selects the optimal parameters using cross-validation.

@@

Detecting edges is very straightforward using an edge detection algorithm like Canny Edge Detection \cite{canny}. However, there might be a lot of edges on the road itself (think of road markings and cobbles). The edges on the road, considered as noise, and the edges of the road are not unambiguously separable. The width of the patterns that cause the edges on the road is rather small. Using morphological image processing operations, it is possible to remove these small patterns.  In figure \ref{cobbles_erode_out}, a smoothed cobbled road is shown.
\begin{figure}[ht]
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\textwidth]{fig/cobbles_in.png}
  \caption{Original frame\label{cobbles_in}}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\textwidth]{fig/cobbles_erode_out.png}
  \caption{Eroded cobbled road\label{cobbles_erode_out}}
\end{subfigure}
\caption{Perform eroding to remove the edges on the road.}
\end{figure}

Eroding \cite{sreedhar2012enhancement} is a morphology operator to make the objects on the foreground, which are the brightest, smaller. Figure \ref{line_zoom_eroded} shows an example of a thinned white line. When the kernel is large enough, white lines can be filtered out completely. The same technique is used to smooth the surface of cobbled roads. A cobblestone consist of a bright center, surrounded by darker joints. By eroding the cobbled road, the erode function will minimize the center of the cobble, which will cause enlarged joints. With a large kernel, the joints will spread until they overlap. The best results were achieved with an ellipsoidal kernel of 100 x 30 px.

@@

Unfortunately, the erode function will cause the darker objects to expand. The original size of the objects has been modified, as seen in \ref{car_erode_bigger}. This will falsify the eventual edge detection. In this example, the observed car is reported too close. In order to restore the original measures, dilation is executed. This is the opposite morphology operator of eroding. The combination of eroding and dilation is called opening. When dilation is executed with the same kernel size of the erode function, the original size of the dark objects will decrease to their original size, as seen in \ref{car_dilate_smaller}. Note that, after eroding, some details of the shape of the objects are lost. In order to provide a safety margin, the dilate function is executed with a slightly smaller kernel.

@@

This algorithm requires a minimum and a maximum threshold value, which were determined experimentally. 
The best results were achieved with a maximum threshold that is twice the minimum threshold.
The minimum threshold is determined by the required minimum sensitivity. When the algorithm is configured too sensitive, some imperfections on the road will be detected as edges. On the other hand, a certain level of sensitivity is necessary to detect road edges that are  less clear. For example, a transition from road to dirt has a very small distinction of color and brightness, as seen in \ref{transition_road_dirt_eroded}. The minimum threshold should be high enough to detect this transition as edge.

@@

@@

In order to calculate the maximum speed a vehicle can have in a particular frame, we use a combination of road classification and edge detection. Like that, we can detect where it is safe to drive. For each frame in the dataset, some masks are given that show the projected trajectory of the car in that particular frame. Intersections between this mask and the ''safe zone'' in the frame get calculated in order to find the maximum safe speed for each frame. The edges of the edge detection are visualized as blue lines on figure \ref{visualization}. The intersections are shown in green. The closest intersection to the car is shown in red. At this point the maximum speed is derived from the intensity of this pixel.

In order to calculate the edges, two different strategies were tested:
\begin{enumerate}
\item  A region is considered to be road only if it's identified as road by the road classification and there are no edges. Both classification and edge detection should indicate a pixel as road in order to be used for safe driving.
\item A region is considered to be road if there are no edges. When edges are detected, road classification is used to decide what part of the image is road. 
\end{enumerate}

Both strategies start by removing outliers from the road classification. The goal of this is to eliminate the effect that outliers would have when it's close to the car. A non-road block is considered an outlier when it has less than 4 non-road neighbors. This is repeated until all outliers are removed.

For the first strategy, the edges of the area classified as road are computed. These get combined with the edges from the edge detection. The closest intersection of any edge with the given mask of the image is used to calculate the maximum speed of the vehicle.

In the second strategy, the edges are first calculated using edge detection. Each edge is then checked whether it lies on the road or not using road classification. For safety reasons there is an additional check. The above, left and right blocks of an edge should also be considered as non-road. When an edge lies on a part of the frame that is not classified as road, it is added to a collection. The intersections of the edges in the collection and the masks are then computed.

@@

@@

@@

The second strategy for combination also results in zero crashes over the four datasets. This strategy tries to focus on edge detection. A region is considered road if there are no edges. When there are edges, road classification is used to decide whether it's road or not. This strategy gave us the best results, as shown in figure \ref{distance_result} as the fourth bar.
\begin{figure}[ht]
	\centering
	\includegraphics[width=\textwidth]{fig/distance_result}
	\caption{Results of road classification, edge detection and combinations.\label{distance_result}}
\end{figure}
